### 2020-09-03 learning Logic/draft.tex
a

\documentclass{article}
\usepackage{graphicx,xcolor}
\usepackage{algorithmicx,algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amssymb,amsmath}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}

\tikzset{
	->, % makes the edges directed
	>=stealth', % makes the arrow heads bold
	node distance=3cm, % specifies the minimum distance between two nodes. Change if necessary.
	every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
	initial text=$ $, % sets the text that appears on the start arrow
}

\newcommand{\jana}[1]{{\color{magenta} #1}}
\newcommand{\kush}[1]{{\color{blue} #1}}

\newcommand{\Real}{\mathbb{R}}
\newcommand{\obs}{\mathrm{obs}}
\newcommand{\free}{\mathrm{free}}
\newcommand{\AP}{\Sigma}

\begin{document}

\section{Problem Setup}

Let $X$ be a bounded open subset of $\Real^n$ and let $X_\obs \subset X$ be the obstacle region. $X_\free = X \setminus X_\obs$ is called the obstacle-free space.

A few more things to define in the standard way:
\begin{itemize}
\item A path
\item A collision-free path
\item A ball
\end{itemize}

Consider a labeling $L: X \rightarrow 2^\AP$ \jana{TODO formalize and assume the labeling induces a partitioning into a finite number of regions}.

Consider a spec given as an \jana{LTL/sc-LTL?} formula over $\AP$.

The problem is to find \jana{a path/an optimal path?} that satisifes the spec. The problem can be solved using straightforward extension to existing algorithms. \jana{Our contribution is a more efficient procedure that ???}

\section{Solution/Contribution draft}

We get a simulation, we project it to the automaton, see what went wrong (is violated or not satisfied yet), in the next try we aim at spotting where the wrong item can be fixed by a different decision.

\begin{itemize}
	\item Learning a small model from the traces, solving that one, lifting the strategy to the original model.
	\item Looking at what remained to be fullfiled and where in the learnt model this is possible.
	\item For unbounded properties, we have to do this hiearchically and check the progress for subformulas. We can do that, using the labelling from Rabinizer.
\end{itemize}

\begin{algorithm}
	\caption{Basic algorithm}\label{alg:general}
	\begin{algorithmic}[1]
		\State build trivial automaton
		\Repeat\Comment{Start new simulation}
		\Repeat\Comment{Next step in simulation}
		\State \Call{ask}{} automaton for advice which Ap to satisfy next
		\State sample a bit to get that one satisfied
		\Until{sampled all the way to satisfaction of the property or too long}
		\State \Call{learn}{} the automaton using also the new sampled paths
		\Until{sampled all the way to satisfaction of the property (fast enough)}

		\Procedure{ask}{}
		\State consider product of the learnt (system) automaton (DFA) and of the property automaton ($\omega$-automaton)
		\State for DFA use shortest path to target
		\State for DBA too
		\State for general ones we might need to non-deterministically choose: which Rabin pair of the DRA? which direction in the NBA?
		\EndProcedure
		\Procedure{learn}{}
		\State automata learning or even probabilistic automata/MDP learning
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\paragraph{Example 1}
Formulae beyond safety: things like $\varphi=GF(a\wedge Xb)\wedge FG(c\wedge dUe)$ do not even have a deterministic B\"uchi automaton, not to mention a DFA, so how to measure progress is not clear form classical approaches. 
(In DFA one can measure the distance to the accepting state, extensible for B\"uchi - not sure this is done, beyond B\"uchi hard.)
Rabinizer tracks the evolution of both parenthesised expressions and after an $a$ this advises us to go for $b$ althought requirement to be satisfied is still exactly $\varphi$.

\begin{figure}[h]
	\includegraphics[scale=0.4]{Fig/auto}\caption{Models for the second automata example}
\end{figure}

\paragraph{Example 2}
Write $\alpha\to\beta$ for $\alpha\wedge F\beta$
Within safety, but taking into account the model: $\psi=F(a\to b\to c)\vee F(d\to e)$.
Classical technique (get closer to the accepting state) prefers $d$ 9get to distance 1) over $a$ (distance still 2). 
However, if we see that nothing reasonable comes after $d$, in particular not even an $a$ then we want to try out the first option, too.
To this end we can build a model from learnt traces to see that taking $d$ does not allow for a winning trace.

Similarly, for $\wedge$ instead of $\vee$, getting $d$ is bad at first (no $a$ after $d$, better to stay on spot than to shorten the distance badly).
Seen orders of subsets of $\{a,b,c,d,e\}$ induce a model.
We can make a product with the automaton and then decide there, where to go.

Jana's notes from the call: 
\begin{itemize}
\item State space; partitioned -- we don't know the labels of regions apriori, but if we have a concrete sample, I am able to test whether an atomic proposition holds or not; make sure we formalize this as an assumption; needed: an atomic proposition does not change infinitely many times during a finite time horizon (zeno formulation??)
\jana{Hmm, how can we not have labels of regions apriori, but be able to test samples at all??}
\item Fast finding of A path that satisfies a complex mission. Pre-processing step to Informed RRT* or something like that. Prove that the important properties are maintained beyond reachability.
\jana{Hmm, how can this further guide informed RRT*? I have logically constrained stuff, so the distance is not the metric to guide it now. What is the metric?}
\end{itemize}

\newcommand{\must}{\Rightarrow}
\newcommand{\may}{\rightarrow}
\newcommand{\maynot}{\nrightarrow}
\newcommand{\mustnot}{\nRightarrow}

\begin{algorithm}
	\caption{one simple instantiation}\label{alg:learn}
	\begin{algorithmic}[1]
		\Require ``multi-modal'' transition system $(S,s_0,\must,\may,\maynot,\mustnot)$, $S\subseteq 2^{Ap}$, transition relations disjsoint (generally any quantity $S\times S\to \mathit{OrderedDomain}$, for instance a counter; here just $\{0,1,2,3\}$)
		\Procedure{learn}{transition from $s$ to $s'$}
		\State $t:=(s_1,s_2):=$(Labelling(s),Labelling(s'))
		\If{$t\notin\must$}
		\State add $t$ to $\must$
		\State domainOf Changes$:=\{a\in Ap\mid s_1(a)\oplus s_2(a)\}$
		\For{$s\in S$ agreeing with $s_1$ on domainOf Changes}
		\State \Comment or any other ``neighbourhood'' criterion / one can learn planning-style rules
		\State \Comment(and under some further conditions so that we don't overwrite $\may,\maynot$ back and forth)
		\State add (s,s changed to $s_2$ on domainOfChanges) to $\may$
		\EndFor
		\EndIf
		\State $counter(s_1)++$
		\If {$counter(s_1)>threshold$}
		\For{$s\in S: (s_1,s)\notin$ any relation}
		\State add $(s_1,s)$ to $\maynot$
		\EndFor
		\EndIf
		\EndProcedure
		\State\Comment and $\mustnot$ requires some really good reason to believe there is no transition
		\Procedure Ask{s}
		\State try using the best $\must$-successor or sample long to get an even better $\may$-successor
		\EndProcedure
		\newline
		\State Comments: 
		\State (1) $\maynot$ can be rather written as a $\may$ into ``?'' from where we can go anywhere; ASK tries to win with as few ?'s as possible
		\State (2) $\mustnot$ could be written as $\must$ into a sink ``:-(''
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Current Implementation: MAIN}\label{alg:current_main}
	\begin{algorithmic}[1]
		\State build trivial automaton
		\State $currentStates=initState$
		\While{true}
			\State $reachableStates=\Call{ask}{currentStates}$ 
			\State \Comment{array of states backward reachable from final states}
			\State $pathLength=1$
			\While{$pathLength<reachableStates.size()$}
				\State $fromStates=currentStates\wedge reachableStates[pathLength]$
				\State $toStates=reachableStates[pathLength-1]$
				\State$t:=(p,q')=\Call{Sample}{fromStates, toStates}$
				\If{$t$ $!=null$}
					\State break
				\EndIf 
				\State $pathLength++$
			\EndWhile
		\If{$t==null$}
		\State $t=\Call{Sample}{currentStates}$
		\EndIf
		\State $currentStates$ $\vee= q$
		\State $\Call{learn}{p,q}$
		\If{accepting path exists}
		\State \Return{$path$}
		\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Current Implementation: LEARN}\label{alg:current_learn}
	\begin{algorithmic}[1]
		\Procedure{learn}{$s_1,s_2$}
		\State $t=s_1\wedge s_2'$
		\State $sampledTransitions\ \vee = t$
		\State $removeTransitions(t)$\Comment{remove transition of smaller level}
		\State $t=\Call{addTransitions}{t,3}$	\Comment{$t$ will be converted to product automaton transitions}

		\State $DOC^c=\{a\in APSystem\mid \neg(s_1(a)\oplus s_2(a))\}$\Comment{Used for $restrict$}
		\State $R=remove(s_1,DOC^c)$\Comment{states smilar to $s_1$}
		\State $newTransitions=R \wedge \neg R' \bigwedge\limits_{a\in DOC^c} a\leftrightarrow a' \wedge level=2$
		\State\Comment{$\neg R':= \bigwedge\limits_{a \text{ in } R} \neg a'$}
		\State $newTransitions$ $\wedge=\neg sampledTransitions$
		\State $productAutomaton $ $\vee= (newTransitions \wedge propertyBDD)$
		\State $counter(s_1)++$
		\If {$counter(s_1)>threshold$}
		\For{$s\in S: (s_1,s) \wedge sampledTransitions = false$}
			\State \Call{removeTransitions}{$s_1,s_s$}
			\State \Call{addTransition} {$s_1,s,1$}
		\EndFor
		\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Current Implementation: ASK}\label{alg:current_ask}
	\begin{algorithmic}[1]
		\Procedure{ask}{$fromStates$}
		\State $finalStates=$\Call{$findFinalStates$}{$productAutomataBDD\wedge fromStates$}
		\State $reachableStates.add(finalStates)$
		\State $reachableStates.add(preFinalStates)$\Comment{required for transition acceptance}
		\State $i=1$
		\While{$\neg allReachableStates \wedge \Call{pre}{reachableStates[i]}\ !=\bot$}
			\State $reachableStates.add(reachableStates[i]\vee \Call{pre}{reachableStates[i]})$
			\State $allReachableStates\ \vee=reachableStates[i]$
			\State $i++$
		\EndWhile
		\State \Return{$reachableStates$}
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\subsection{LEARN and ASK procedures}

\textit{LEARN:} Learn procedure takes a transition $t=(s_1, s_2)$ (or a set of transitions) and add similar `maybe' transitions to the product automaton. For defining these similar transitions we first need a set of atomic propositions called domain of changes, which is defined as $\{a \oplus  b \mid a\in s_1, b\in _2\}$.  A transition similar to $t$ would be $(s_1', s_2')$ where $s_1'\ (s_2')$ is a state which agrees with $s_1\ (s_2)$ on domain of changes. For e.g, if $t = (\{r_1\}, \{r_1, b\})$ then, the set $DOC=\{b_1\}$ and one of the learned transition would be $(\{r_2\},\{r_2, b\})$. If we have a different atomic proposition for all the bins and $t = (\{r_1\}, \{r_1, b_1\})$ then the learned transition will be $(\{r_2\}, \{r_2,b_1\})$ [not $(\{r_2\}, \{r_2,b_2\})$].

\textit{ASK:} Ask procedure takes the product automaton and returns a list of set of states called $reachableStates$ whose $i$th element is a set of states which can see an accepting transition in the product automaton in $\leq i$ steps. Look at the example in figure \ref{fig:ask}, transition with label $(*)$ is an accepting transition. $reachableStates[0]=\{s_6\}$, [1]=$\{s_5, s_6\}$, [2]=$\{s_3,s_5,s_6\}$ and [3]=$\{s_0,s_1,s_2,s_3,s_5,s_6\}$

\begin{figure}
	\begin{tikzpicture}[
	roundnode/.style={circle, draw=black, fill=black!5, very thick, minimum size=3mm}]

	\node[roundnode] (A) {$s_0$};
	\node[roundnode] (B) [below=of A] {$s_1$};
	\node[roundnode] (C) [below=of B] {$s_2$};
	\node[roundnode] (D) [right=of A] {$s_3$};
	\node[roundnode] (E) [below=of D] {$s_4$};
	\node[roundnode] (F) [right=of D] {$s_5$};
	\node[roundnode] (G) [right=of F] {$s_6$};

	\draw[->] (A) -- (D);
	\draw[->] (B) -- (D);
	\draw[->] (C) -- (E);
	\draw[->] (C) -- (D);
	\draw[->] (D) -- (F);
	\draw[->] (F) -- (G) node[anchor = south, pos = 0.5] {$(*)$};
	\end{tikzpicture}
	\caption{A part of the product automaton}
	\label{fig:ask}
\end{figure}

\subsection{A mini case study}

Let us consider an office space with several rooms as illustrated in Fig. \ref{fig:office}. There are different objects in the office rooms. For simplicity, let us assume that each of the objects is either a desk (illustrated as a rectangle), or a chair (illustrated as a square), or a waste bin (illustrated as a circle). Consider a fully actuated robot with the initial position $x_0 \in \mathbb{R}^2$. We assume that the robot can recognize if it is close to a certain type of object or landmark in the environment; by close, we mean that the robot's state intersect with the object's/landmark's $\epsilon$-inflation (illustrated as dashed and dotted lines).

The set of atomic propositions is as follows:
\begin{itemize}
\item $\mathit{Room}_1,\ldots,\mathit{Room}_6, \mathit{Hall}$, stand for being in a particular room or in a hallway
\item $\mathit{Desk}$ stands for being close to a desk
\item $\mathit{Chair}$ stands for being close to a chair
\item $\mathit{Bin}$ stands for being close to a bin
\item $\mathit{Corner}$ stands for being close to a corner
\item $\mathit{Wall}$ stands for being close to a wall
\end{itemize}

The specification for the robot's behavior is given by the LTL formula:
$$G F (\mathit{Room}_1 \wedge \mathit{Bin}) \wedge \ldots \wedge G F (\mathit{Room}_6 \wedge \mathit{Bin}).$$

\begin{figure}
\centering
\includegraphics[width=\linewidth]{Fig/small-office} \caption{A mini case study office environment illustration.}
\label{fig:office}
\end{figure}

\subsection{A sample run of the algorithm}
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{Fig/example} \caption{A smaller example to illustrate the algorithm}
	\label{fig:example}
\end{figure}

Look at the system in Figure \ref{fig:example}. The property we want to satisfy for this system is $G F (r_1 \wedge b) \wedge G F (r_2\wedge b)$ and the deterministic B\"uchi automaton for this LTL formula is also given in Figure \ref{fig:example}. The algorithm stores the product automaton as a BDD, intially it is $\bot$. $currentStates$ stores the set of states we have seen till now which is initialized to $initState$. For this Example $initState=\{H,0\}$.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{Fig/product2} \caption{Product automaton after 2 iterations}
	\label{fig:product2}
\end{figure}

After sampling transitions $\{H\} \rightarrow \{r_1\}$ and $\{r_1\}\rightarrow \{r_1,b\}$ initially, the product automaton looks like figure \ref{fig:product2}. Now, in line 4, $ASK$ procedure returns a list of BDDs such that $i$th indexed element is a set of states from which there exist a path in the product automaton of length $i$ whose last transition is an accepting one. In lines 7-13, it tries to sample a transition. Initially it will try to sample a transition from states which are at distance 1 from the accepting transitions and then it will increase the length with each iteration. The procedure $SAMPLE$ on line 10 will sample transitions until it finds a transition from $fromStates$ to $toStates$ or it exceeds a given threshold time/number.
If it is still not able to find any "good" transition then just sample randomly from any one of the state in $currentStates$ (line 14-15). Suppose it samples $\{r_1\}\rightarrow \{H\}$ next. Line 16 adds post of the sampled transition i.e $\{H\}$ into $currentSates$. Line 17 adds the sampled transition of level 3 in the product automaton and a few other "similar" transitions of level 2. For our example the product automaton after 3rd iteration look like figure \ref{fig:product3}.

For the example given in Figure \ref{fig:office}, without considering $wall$ atomic proposition, it was taking approx. 15 seconds to do everything. There is some room for improvement and it might get reduced to approx. 10 seconds.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{Fig/product3} \caption{Product automaton after 3 iterations}
	\label{fig:product3}
\end{figure}

\vspace{20pt}
\textbf{Points from the discussion on 13.05.2020:}
	\begin{itemize}
		\item We assume that the robot has the labelling of the whole system beforehand for now. We can consider the other case later.

		\item Use RRG instead of RRT so that we can go back also.

		\item \textbf{What we need from the SAMPLE procedure:} Given a pair of set of states (fromStates, toStates) in the abstract system (formulas over the APs of the system), SAMPLE should return a newly sampled transition from a state where `fromState' is \textit{true} to a state where `toState' is \textit{true}. It maybe the case that such a transition doesn't exist or probability of it being sampled is very small, in that case, return \emph{null} after some given threshold time/samples.\newline
		\textbf{How do we do this with RRG?}
		\begin{itemize}
			\item Each node in the graph ($G$) will store the APs associated with it. 
			\item If the whole tree is stored as a BDD, finding nodes where `fromStates' is very easy but storing it in a BDD is very difficult. \kush{Find a way to do both the things efficiently.}
		\end{itemize}

		\item \textbf{Jan's idea:} We can keep track of the sampled transitions till now and get their ranking (finding ranking for each sampled transition will require some time because it might have to pass through the whole list of `reachableStates'). Sample 50 (some threshold) transitions and add the one with the best rank. \kush{What will happen to the other transitions? Would we keep all of them in the memory (maintain a min-heap maybe)?}
	\end{itemize}

\section{Fernando's Ideas}
Suppose the system knows the environment, it's free space and obstacles. I.e. $X_{\text{obs}}, X_{\text{free}}$ are known. Let's also say that the set of $k$ atomic propositions $Ap$, denoted as $\Sigma = \{Ap_1, Ap_2, \dots, Ap_k\}$, is known. However, let's suppose the labeling function $L: X \to 2^\Sigma$ is partially known, and constructed as the robot navigates the environment. For example, it can attribute atomic propositions to regions of the workspace when within $\epsilon$-distance to it.

I think approaching it this way would make our contribution richer. If we assume to know $L$, we actually don't need to know anything else about the workspace besides the positions of $Ap$s in question, for example the bins. No spatial relation is necessary. Now, if $L$ is not fully known, the problem becomes more interesting. Our system will then try to learn and predict characteristics of the workspace as it moves through it.

Also, what comes to my mind is regarding the mission formula and how much information we feed to it. In $G F (r_1 \wedge b) \wedge G F (r_2\wedge b)$ we are already giving out that there exists a bin in each room. If one of the rooms does not have a bin, the formula is not satisfiable anymore. If the goal is to collect all bins, the formula should somehow contain only bins. The relation of bins being inside rooms should be learned by the model.

Now, how to make sure the robot does not visit the same bin infinitely often? Should we know beforehand the number of bins and write a formula like $\bigwedge_{i \in [1,n]} GF b_i$?

This raises another question: is there any grouping of $Ap$s? How does the algorithm knows that $r_1$ and $r_2$ are both rooms, and not a room and a lake? Or the inference that a transition from $r_1$ to $b_1$ can mean that there might exist a transition from $r_2$ to $b_2$, but not from $h$ to $b_3$?

\end{document}