### 2021-02-24 Presentation/Jana's group/scheme.tex
a

\begin{figure}
	\centering
    \begin{tikzpicture}[->,>=stealth', minimum width = 32mm,text width = 30mm]
        \node[state=initial, color = blue, anchor=center, fill = blue!10!white,thin] (A) {System};
        \node[state, color = blue, below of = A, fill = blue!10!white, node distance =1.5cm,thin] (B) {RRG graph};
        \node[state, color = blue, below of = B, fill = blue!10!white, node distance =1.5cm,ultra thick] (C) {Semantic abstraction};
        \node[state, color = red, below of = C, fill = red!10!white, xshift = -2cm, node distance =1.5cm,thin] (D) {Product};
        \node[state, color = green!50!black, left of = A, fill = green!10!white, node distance =4cm,thin] (E) {LTL property};
        \node[state, color = green!50!black, below of = E, fill = green!10!white, node distance =2.25cm,thin] (F) {Automaton};

        \path (A) edge[] (B)
        (B) edge[ultra thick] node[right,pos=0.5]{\texttt{Learn}} (C)
        (C) edge (D)
        (E) edge (F)
        (F) edge (D)
        (D.east) edge[bend right=70,looseness=1.5, ultra thick] node[right,pos=0.9]{\texttt{Bias}} (B.east)
        %guidance $\rightarrow$ better paths and learning w.r.t. property
        ;
    \end{tikzpicture}
    \caption{Scheme of our model-checking-inspired approach with novel elements drawn thickly. }
    \label{fig:guide}
\end{figure}